---
layout: post
description: "Working on visualization on R using the ggplot library"
title: "A quick analysis on the brazilian profile of the candidates for election"
categories:
  - Data Analysis
tags: [Unix,R,Visualization]  
---

### Visualization with R

This post will be a simple analysis on electoral data from Brazil. I'll explore some points at the data and draw beautiful plots using **ggplot** package. I hope you enjoy the potential of R to make informative and clean graphs.
  
```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(data.table)
library(magrittr)

library(ggplot2)
library(scales)
library(extrafont)

library(RCurl)
```

A brief explain on the packages:

  - Dplyr and data.table are the undispensible tools for every analysis. 
  - GGplot2, scales and extrafont are the best tools for static plots.
  - Rcurl is good workaround package to solve download issues.
  
Before going further, let's get a single file and have a overall look on it. The candidates file of electoral race can be downloaded from the Superior Electoral Court website. 

```{r first-look}
tf <- tempfile(); td <- tempdir()
url = 'http://agencia.tse.jus.br/estatistica/sead/odsele/consulta_cand/consulta_cand_2016.zip'
download.file(url, destfile = tf)
unzip(tf, exdir = td)
```

The *system* R function calls a bash command. Let's use them to have a look on the files. The **system** function have a different way of outputting the results. This is a problem that can easily be overcome by a little hacking.

```{r bash-commands}
# To capture the bash output
system = function(...) cat(base::system(..., intern = TRUE), sep = "\n")
# Filesize
system(sprintf('cd %s; wc -c consulta_cand_2016_SP.txt', td))
# Head
system(sprintf('cd %s ; 
head -n 3 consulta_cand_2016_SP.txt | iconv -f iso-8859-1 -t UTF-8', td))
```

The, likely, biggest file have only 46 Mb.
The file does not have a header, the *NULL* entries are **"#NULO#"** and the separator is **";"**. The year states files come with a README pdf. There you can see the header. 

The next step is do it in a loop so I can have candidates files candidates file for the 2004 year until 2016. As the files comes separeted by states, we have to bind them into one. It can be done in R, but it's easier and faster in **bash** using *cat*.

Fortunately I found a Github repository that have mapped the header - [Github silvadenisson](https://github.com/silvadenisson/electionsBR/blob/master/R/candidate_local.R). I saved the naming part of the file in my Github. Know, let's use it.

#### Downloading and transforming the raw data

```{r loop-read}
candidates_data <- list()
# Load the function that brings the correct header
eval(expr = 
       parse(text = getURL("https://raw.githubusercontent.com/MatheusRabetti/matheusrabetti.github.io/master/assets/posts/candidates-beautiful-plot/header.R",
                           ssl.verifypeer=FALSE) ))

for(ano in seq(2004, 2016, 4)){
  cat('Downloading year', ano, '\n')
  
  # Download and save in a temporary folder
  sprintf('http://agencia.tse.jus.br/estatistica/sead/odsele/consulta_cand/consulta_cand_%s.zip',ano) %>% 
    download.file(destfile = tf, quiet = T)
  unzip(tf, exdir = td)
  
  cat('Binding and reading the files ...\n')
  # Binding
  system(sprintf('cd %s; cat consulta*%s*.txt >> all-%s.txt', td, ano, ano))
  # Reading
  candidates_data[[as.character(ano)]] <- 
    fread(sprintf('%s/all-%s.txt',td,ano), 
          header = F, sep = ";", 
          stringsAsFactors = F, 
          encoding = "Latin-1")
  
 
  candidates_header(candidates_data[[as.character(ano)]])
  
}


candidates_data <- data.table::rbindlist(candidates_data, use.names = T, fill = T)

```

I'll select the variables listed below to continue my analysis:

  - **ANO_ELEICAO:** Election year
  - **DESCRICAO_CARGO:** Description of the position that the candidate runs for.
  - **DESCRICAO_OCUPACAO:** Candidate's occupation description.
  - **CODIGO_SEXO:** Candidate's sex code.
  - **DESCRICAO_SEXO:** Candidate's sex description.
  
At first I want to see the percentage of women that canditates in politics in Brazil. So, let's use **dplyr** to manipulte the data.  


```{r select}
table(candidates_data$CODIGO_SEXO)

candidates_data %>% 
  select(ANO_ELEICAO, DESCRICAO_CARGO, DESCRICAO_OCUPACAO, 
         CODIGO_SEXO, DESCRICAO_SEXO) %>% 
  filter(CODIGO_SEXO %in% c(2,4)) %>%
  group_by(ANO_ELEICAO) %>%
  summarise(fem_percent = round(sum(CODIGO_SEXO == '4')/n(), 4)) %>% 
  mutate(xtralabs = ifelse(fem_percent == max(fem_percent), 'Highest:\n', 
                           ifelse(fem_percent == min(fem_percent), 'Lowest:\n',''))) -> female_percent
knitr::kable(female_percent)
```

#### Analysis - drawing a portrait of electoral candidates

Before you run the next command I advise you install de [extrafont package](https://github.com/wch/extrafont), indispensable if you use Linux. The font you use in a plot absolutely changes the view of it. This package helps you administrates your fonts and even install new.

```{r women-percentage}
female_percent %>% 
  ggplot(., aes(ANO_ELEICAO, fem_percent)) +
  geom_bar(stat = 'identity', width = .65) +
  geom_label(aes(label = sprintf('%s%s', xtralabs, percent(fem_percent))),
             vjust = -.4,
             family= rep("Verdana", 4),
             fontface = c('bold', rep('plain',3)),
             lineheight=.9, size=4, label.size = 0) +
  scale_x_discrete() +
  scale_y_continuous(expand = c(0,0), limits = c(0.0,1.0), labels = percent) +
  labs(x=NULL, y=NULL, title="Portrait of Brazilian Politics",
       captions = 'Data from TSE/Repositório Eleitoral',
       subtitle = 'Female percentage of electoral candidacy') +
  theme_minimal(base_family = 'Verdana') +
  theme(axis.line = element_line(color='#2b2b2b', size =.5) ) +
  theme(axis.line.y = element_blank()) +
  theme(axis.text.x = element_text(family = rep("Verdana", 4),
                                   face = c('bold', rep('plain', 3)))) +
  theme(panel.grid.major.x=element_blank()) +
  theme(panel.grid.major.y=element_line(color="#b2b2b2", size=0.1)) +
  theme(panel.grid.minor.y=element_blank())+
  theme(plot.title=element_text(hjust=0, 
                                family="Verdana",
                                face = 'bold',
                                margin=margin(b=10))) +
  theme(plot.caption = element_text(hjust = 0,
                                    family = 'Verdana',
                                    margin = margin(t = 20))) + 
  theme(plot.subtitle = element_text(margin = margin(b = 20),
                                     family = 'Verdana'))
```

The situation is getting better, but is far from being considered good. There is still a gap of 18.1% to the ideal cenary of half of candidates.

